### 2021-11-18 Complexity/UGC/approx.tex
First draft of UGC paper
SDP and Goemans written

\begin{example} There exists an Semi-definite Programming (SDP) relaxation for $\mathsf{MaxCut}$.

>>>>

\begin{example} \label{maxcutexample}
The problem $\MaxCut$ will be casted as follows: given an undirected graph $G=(V,E)$, find the largest cut of $G$. There is a straightforward LP formulation for this problem:

\begin{align}
  &  \min \sum_{(u,v) \in E} e_{u,v} \\
  & e_{u,v} \geq
  \max \begin{cases}
      x_1 + x_2 \\
      2 - (x_1 + x_2)
  \end{cases} \\
  & e_{u,v} \in [1,2] \\
  & x_v \in [0,1]
\end{align}
%
where the $e_{u,v}$ variables represent the edges $e \in E$ of the input graph and the $x_v$ represent the vertex variables $v \in V$. This will actually be a relaxation from its respective integer program in which the variables can take the integer values:

\begin{align}
  e_{u,v} = \begin{cases}
              1 \text{ if } e_{u,v} \in \mathcal{C} \\
              2 \text{ otherwise }
            \end{cases}
  x_v = \begin{cases}
              0 \text{ if } v \text{ is in partition } S \\
              1 \text{ otherwise }
        \end{cases}
\end{align}

where $\mathcal{C} \subseteq E$ is the edges constituting the cut and $(S,\bar{S})$ is the partition defining the cut. Naturally, a solution to the Integer program will also be a solution to the LP above, which implies that

\[ \mathsf{Opt}_{LP} \geq \MaxCut(G)\]

There actually exists an Semi-definite Programming (SDP) relaxation for $\MaxCut$ called the \emph{Goemans-Williamson} algorithm which will be treated in Section \ref{goemans}.

### 2021-11-18 Complexity/UGC/labelcover.tex
SDP and Goemans written

\section{Label-Cover and Projection Games}
%Definition of Label-Cover
We now introduce a problem which manages to provide a natural paradigm for capturing the essence of CSPs and proving inapproximability results. These ``projection games" were introduced by Bellare, Goldreich, and Sudan \cite{bellare1998free}. The $NP$-hardness of the gap problem version of Label Cover was used by H\aa stad to show tight inapproximability results for $\mathsf{Max3SAT}$ and $\ELin$ \cite{haastad2001some}.

@@

%\subsection{}
%Arora, Barak
\subsection{Raz's Parallel Repetition Theorem}
To be written...

>>>>

\section{Label-Cover and Projection Games}
%Definition of Label-Cover
We now introduce a problem which manages to provide a natural paradigm for capturing the essence of CSPs and proving inapproximability results. These ``projection games" were introduced by Bellare, Goldreich, and Sudan \cite{bellare1998free}. The $\NP$-hardness of the gap problem version of Label Cover was used by H\aa stad to show tight inapproximability results for $\mathsf{Max3SAT}$ and $\ELin$ \cite{haastad2001some}.

@@

%\subsection{}
%Arora, Barak
\subsection{Some Structural Results of PCPs}

\subsection{Raz's Parallel Repetition Theorem}

### 2021-11-18 Complexity/UGC/ugcmaxcut.tex
SDP and Goemans written

\newcommand{\GapMaxCut}{\mathsf{Gap}_{\GWConstant + \epsilon, \frac{1-\rho}{2} - \epsilon}\mathsf{MaxCut}}
\newcommand{\ULC}{\mathsf{UniqueLC}(m)}
\newcommand{\GapULC}{\mathsf{GapUniqueLC}(m)}
\newcommand{\GapDeltaULC}{\mathsf{Gap}_{\delta, 1-\delta}\ULC}

@@

\subsection{Goemans-Williamson Algorithm for $\MaxCut$}
Recall that Example \ref{maxcutexample} presented an LP-based approximation algorithm for $\MaxCut$. Goemans and Williamson \cite{goemans1995improved} designed an SDP to give an $\alpha_{GW}$-approximation algorithm for $\MaxCut$ where:

@@

@@

\begin{theorem} \label{maxcutughard} ($\MaxCut$ is $\UG$-hard) The problem $\GapMaxCut$ is $\UG$-hard]
\end{theorem}
Note that an immediate corollary of Theorem and the UGC is that the Goemans-Williamson algorithm is tight:

>>>>

@@

\subsection{Goemans-Williamson Algorithm for $\MaxCut$} \label{goemans}
 Example \ref{maxcutexample} presented an LP-based approximation algorithm for $\MaxCut$. Goemans and Williamson \cite{goemans1995improved} designed an SDP to give an $\alpha_{GW}$-approximation algorithm for $\MaxCut$ where:

@@

To begin, a semi-definite program is a generalization of a linear program where instead of optimizing over a vector of variables $\vec{x}$, the program considers a positive semi-definite matrix of variables i.e a matrix whose eigenvalues are non-negative. An equivalent formulation considers inner products between pairs of vectors:

\begin{align*}
     \max & \sum_{i,j} c_{ij}\langle v_i, v_j \rangle \\
     \text{ under constraints }& a_{ij}^k\langle v_i, v_j \rangle \leq b_{ij}^k \\
     v_1, \cdots v_n \in \mathbb{R}, \quad & k \in [C] \text{ constraints }
\end{align*}

Note that this form also subsumes quadratic programming by setting $c_{ij} = a_{ij}^k = b_{ij}^k = 0$ for all $i \neq j$ and $k \in [C]$. The Goemans-Williamson algorithm concerns the solution to the semi-definite program given some graph $G = (V,E)$: \newline
%
\begin{align}
  & \max \sum_{i,j} \frac{1 - \langle v_i, v_j \rangle}{2} \\
  & \langle v_i, v_i \rangle = 1 \text{ for all } i \in V
\end{align}

Let us first show that indeed the program is a relaxation of the integer program crafted in Example \ref{maxcutexample}. Indeed, if we set any unit vector $\vec{u}$ such that for a cut defined by $(S,\bar{S})$:
%
\begin{align*}
  \vec{v_i} = \begin{cases}
                u \text{ if } i \in S \\
                -u \text{ if } i \not\in S
              \end{cases}
\end{align*}

A direct calculation yields that the term $\frac{1 - \langle v_i, v_j \rangle}{2} = 0$ if $v_i = v_j = u$ else it is equal to 1. Thus, by applying this observation to the maximum cut,
%
 \[\mathsf{Opt}_{GW}(G) \geq \MaxCut(G)\]

So far the semi-definite program seems to be rather simple. The insight made by Goemans and Williamson lies in the rounding procedure of the solution outputted by the program. The procedure proceeds by first drawing a random hyperplane passing through the origin and taking the two partitions of the cut $S_+, S_-$ to be the solution $v_i$ which lie on positive and negative sides of the hyperplane respectively. We can calculate the expected size of the cut:
%
\begin{align*}
\mathbb{E}[|E(S_+,S_-)|] = \sum_{(i,j) \in E} \mathbb{P}[v_i,v_j \text{ lie on different sides of the hyperplane}]
\end{align*}

Now if $\theta_{ij}$ denotes the angle between $v_i,v_j$, a simple geometric argument shows that:

\[\mathbb{P}[v_i,v_j \text{ lie on different sides of the hyperplane}] = \frac{\theta_{ij}}{\pi} \]

By definition of the dot product:

\[\frac{1 - \langle v_i, v_j \rangle}{2} =  \frac{1 - \cos(\theta_{ij})}{2} \]

and the magical inequality:

\[ \frac{\theta_{ij}}{\pi} \geq \alpha_{GW} \cdot \frac{1 - \cos(\theta_{ij})}{2}\]

the two can be combined to finally yield that:

\begin{equation}
  \frac{\mathbb{E}[|E(S_+,S_-)|]}{\mathsf{Opt}_{GW}(f)} \approx \alpha_{GW}
\end{equation}

@@

\begin{theorem} \label{maxcutughard} ($\MaxCut$ is $\UG$-hard) The problem $\GapMaxCut$ is $\UG$-hard
\end{theorem}
Note that an immediate corollary of Theorem and the UGC is that the Goemans-Williamson algorithm is tight:

### 2021-11-18 Complexity/UGC/uniquegames.tex
SDP and Goemans written

\begin{remark}
  The $1-\delta$ constant is crucial to the validity of the conjecture. For an instance of $\mathsf{UniqueLC}(m)$ for any $m$ with a guaranteed solution, there exists a polynomial-time algorithm finding an assignment which satisfies all projection constraints: Start with a vertex and set it to a label. If it's a part of an edge, follow the edge to the other side and find a label which satisfies as many edges as possible. Repeat in a breadth-first search fashion. In virtue of the projection maps being permutations, this amounts to searching for the guaranteed solution in time $O(mn^2)$.
\end{remark}

@@

\begin{definition}
  A problem is said to be \emph{$\UG$-hard}
\end{definition}

>>>>

\begin{remark}
  The $1-\delta$ constant is crucial to the validity of the conjecture. For an instance of $\mathsf{UniqueLC}(m)$ for any $m$ with a guaranteed solution, there exists a polynomial-time algorithm finding an assignment which satisfies all projection constraints: Start with a vertex and set it to a label. If it is an endpoint of an edge $e$, follow $e$ to the other side and find a label which satisfies as many neighbors as possible. Repeat in a breadth-first search fashion. In virtue of the projection maps being permutations, this amounts to searching for the guaranteed solution in time $O(mn^2)$.
\end{remark}

@@

\begin{definition}
  A problem $P$ is said to be \emph{$\UG$-hard} if there is some constant $\delta > 0$ such that for all $m \in \mathbb{N}$, there exists a polynomial-time reduction from $\GapDeltaULC$ to $P$.
\end{definition}
